Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-07-25 draft.tex
Fix issues in Apps. C and D found by Fred

The class of $P\to P_1 P_2 \ell_1\bar\ell_2$ decays \fred{title has $\bar{B}$ instead of $P$}, where $P,P_1$, and $P_2$ are pseudoscalar mesons,
and $\ell_1$ and $\bar\ell_2$ are charged or neutral leptons, includes decays
such as $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$.
Their decay PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, k^2, \cos\theta_1, \cos \theta_2, \phi) = \frac{1}{\langle \Gamma\rangle}\frac{\rmdx{^5 \Gamma}}{\rmdx{q^2} \rmdx{k^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\phi}} = \frac{1}{4\pi} \sum_i \frac{I_i(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\cos\theta_1,\phi)\,,
\end{equation}
\fred{rather use index $k$ as in previous appendix} where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_3 & = \sin^2\theta_1 \cos 2\phi\,,\\
    f_4 & = \sin 2\theta_1 \cos  \phi\,,&
    f_5 & = \sin  \theta_1 \cos  \phi\,,&
    f_6 & = \cos  \theta_1\,, \\
    f_7 & = \sin  \theta_1 \sin  \phi\,,&
    f_8 & = \sin 2\theta_1 \sin  \phi\,,&
    f_9 & = \sin^2\theta_1 \sin 2\phi\,.\\
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $k^2$ and $\theta_2$. The
azimuthal angle is denoted as $\phi$. \fred{Danny changed $\phi$ to $\theta_3$ in previous appendix. Here, too?}
The total decay width is \fred{integrand independent of $\theta_1$? I guess you want to integrate over $\theta_2$}
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \rmdx{\theta_1} \rmdx{q^2} \rmdx{k^2} \sin\theta_2 \frac{3I_1(q^2, k^2, \cos\theta_2) - I_2(q^2, k^2, \cos\theta_2)}{3}\,.
\end{equation}
As usual, we transform from a $\cos\theta_1$-dependence to a $\theta_1$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, k^2, \theta_1, \theta_2, \phi)
    & = \frac{1}{4\pi} \sum_i \frac{I(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \phi) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_i S_i(q^2, k^2, \theta_2) f_i(\theta_1, \phi)\,.
\end{aligned}
\end{equation}
where we also define $S_i(q^2, k^2, \theta_2) \equiv \sin\theta_2 I_i(q^2, k^2, \cos\theta_2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.
\fred{$\sin\theta_1$ missing after $f_i$? Why $\rmdx{\Gamma} /\, \rmdx{q^2}$ and not just $\Gamma$ in definition?}\\

The dual basis reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \hat{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \hat{f}_3 & = \frac{15}{4}                \cos 2\phi     \,,\\
    \hat{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \phi     \,, &
    \hat{f}_5 & = 3            \sin  \theta_1 \cos  \phi     \,, &
    \hat{f}_6 & = 3 \cos\theta_1                   \,,\\
    \hat{f}_7 & = 3            \sin  \theta_1 \sin  \phi     \,, &
    \hat{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \phi     \,, &
    \hat{f}_9 & = \frac{15}{4}                \sin 2\phi     \,.\\
\end{aligned}
\end{equation}
\fred{$\hat{f}_3$ and $\hat{f}_9$ need factor $\frac{4}{5}$}
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \int_0^{2\pi} \rmdx{\phi} \hat{f}_i(\theta_1,\phi) P(q^2, k^2, \theta_1,\theta_2,\phi) = S_i(q^2, k^2, \theta_2)\,.
\end{equation}

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts -- reads \cite{Boeer:2014xx}
\begin{equation}
    P_{\Lambda_b}(q^2, \cos\theta_1, \cos\theta_2, \phi) = \frac{1}{\rmdx{\Gamma} /\, \rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\phi}} = \frac{3}{8\pi} \sum_i \frac{K_{i}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_i(\cos\theta_1, \cos\theta_2, \phi)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ \fred{Then use $\theta_3$ below and not $\phi$} denotes the azimuthal angle.
The basis of angular functions reads \cite{Boeer:2014xx}
\begin{equation}
    \begin{aligned}
% 1
        f_{1ss} & = \sin^2\theta_1 &
        f_{1cc} & = \cos^2\theta_1 &
        f_{1c}  & = \cos\theta_1\\
% 2
        f_{2ss} & = \sin^2\theta_1 \cos\theta_2 &
        f_{2cc} & = \cos^2\theta_1 \cos\theta_2 &
        f_{2c}  & = \cos\theta_1   \cos\theta_2 \\
% 3
        f_{3sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        f_{3s}  & = \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        f_{4sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        f_{4s}  & = \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}
The decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of the cosines, the PDF reads
\begin{equation}
    P_{\Lambda_b}(q^2, \theta_1, \theta_2, \phi) = \frac{3}{8\pi} \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \phi) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_i(q^2) = K_i(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\

The dual basis reads
\begin{equation}
    \begin{aligned}
% 1
        \hat f_{1ss} & = \frac{\pi^3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) &
        \hat f_{1cc} & = \frac{\pi^3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) &
        \hat f_{1c}  & = \pi^3 \cos\theta_1\\
% 2
        \hat f_{2ss} & = \frac{3\pi^3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2cc} & = \frac{3\pi^3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2c}  & = 3\pi^3 \cos\theta_1   \cos\theta_2 \\
% 3
        \hat f_{3sc} & = \frac{15\pi^3}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        \hat f_{3s}  & = \frac{3\pi^3}{2} \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        \hat f_{4sc} & = \frac{15\pi^3}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        \hat f_{4s}  & = \frac{3\pi^3}{2} \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}
such that \fred{Works out only if $\pi^3$ dropped everywhere, and $\hat f_{2cc} \to 2 \hat f_{2cc}$}
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \int_0^\pi \rmdx{\theta_2} \int_0^{2\pi} \rmdx{\phi} P(q^2, \theta_1, \theta_2, \phi) \hat{f}_i(\theta_1, \theta_2, \phi) = S_i(q^2)\,.
\end{equation}

>>>>

The class of semileptonic $\bar{B}\to P_1 P_2 \ell_1\bar\ell_2$ decays, where $P_1$, and $P_2$ are pseudoscalar mesons,
and $\ell_1$ and $\bar\ell_2$ are charged or neutral leptons, includes decays
such as $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$.
Their decay PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, p^2, \cos\theta_1, \cos \theta_2, \theta_3) = \frac{1}{\langle \Gamma\rangle}\frac{\rmdx{^5 \Gamma}}{\rmdx{q^2} \rmdx{p^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\phi}} = \frac{1}{4\pi} \sum_k \frac{I_k(q^2, p^2, \cos\theta_2)}{\Gamma} f_k(\cos\theta_1,\phi)\,,
\end{equation}
where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_3 & = \sin^2\theta_1 \cos 2\phi\,,\\
    f_4 & = \sin 2\theta_1 \cos  \phi\,,&
    f_5 & = \sin  \theta_1 \cos  \phi\,,&
    f_6 & = \cos  \theta_1\,, \\
    f_7 & = \sin  \theta_1 \sin  \phi\,,&
    f_8 & = \sin 2\theta_1 \sin  \phi\,,&
    f_9 & = \sin^2\theta_1 \sin 2\phi\,.\\
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $p^2$ and $\theta_2$. The
azimuthal angle is denoted as $\theta_3$.
The total decay width is
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \rmdx{\theta_2} \rmdx{q^2} \rmdx{p^2} \sin\theta_2 \frac{3I_1(q^2, p^2, \cos\theta_2) - I_2(q^2, p^2, \cos\theta_2)}{3}\,.
\end{equation}
As previous, we transform from a $\cos\theta_2$-dependence to a $\theta_2$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, p^2, \theta_1, \theta_2, \theta_3)
    & = \frac{1}{4\pi} \sum_k \frac{I_k(q^2, p^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \theta_3) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_k S_k(q^2, p^2, \theta_2) f_k(\theta_1, \theta_3) \sin\theta_1\,.
\end{aligned}
\end{equation}
where we also define $S_k(q^2, p^2, \theta_2) \equiv \sin\theta_2 I_k(q^2, p^2, \cos\theta_2) / \langle \Gamma\rangle$.\\

The dual basis reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \hat{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \hat{f}_3 & = \frac{15}{4} \sin^2\theta_1 \cos 2\phi     \,,\\
    \hat{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \phi     \,, &
    \hat{f}_5 & = 3            \sin  \theta_1 \cos  \phi     \,, &
    \hat{f}_6 & = 3 \cos\theta_1                   \,,\\
    \hat{f}_7 & = 3            \sin  \theta_1 \sin  \phi     \,, &
    \hat{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \phi     \,, &
    \hat{f}_9 & = \frac{15}{4} \sin^2\theta_1 \sin 2\phi     \,.\\
\end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \int_0^{2\pi} \rmdx{\phi} \hat{f}_i(\theta_1,\phi) P(q^2, k^2, \theta_1,\theta_2,\phi) = S_i(q^2, k^2, \theta_2)\,.
\end{equation}

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts -- reads \cite{Boeer:2014xx}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \phi) = \frac{1}{\rmdx{\Gamma} /\, \rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\phi}} = \frac{3}{8\pi} \sum_i \frac{K_{i}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_i(\cos\theta_1, \cos\theta_2, \phi)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The basis of angular functions reads \cite{Boeer:2014xx}
\begin{equation}
    \begin{aligned}
% 1
        f_{1ss} & = \sin^2\theta_1 &
        f_{1cc} & = \cos^2\theta_1 &
        f_{1c}  & = \cos\theta_1\\
% 2
        f_{2ss} & = \sin^2\theta_1 \cos\theta_2 &
        f_{2cc} & = \cos^2\theta_1 \cos\theta_2 &
        f_{2c}  & = \cos\theta_1   \cos\theta_2 \\
% 3
        f_{3sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\theta_3 &
        f_{3s}  & = \sin\theta_1              \sin\theta_2 \cos\theta_3 \\
% 4
        f_{4sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\theta_3 &
        f_{4s}  & = \sin\theta_1              \sin\theta_2 \sin\theta_3
    \end{aligned}
\end{equation}
The decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of the cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_i(q^2) \equiv K_i(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\

The dual basis reads
\begin{equation}
    \begin{aligned}
% 1
        \hat f_{1ss} & = \frac{1}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) &
        \hat f_{1cc} & = \frac{1}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) &
        \hat f_{1c}  & = \cos\theta_1\\
% 2
        \hat f_{2ss} & = \frac{3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2cc} & = \frac{3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2c}  & = 3 \cos\theta_1   \cos\theta_2 \\
% 3
        \hat f_{3sc} & = \frac{15}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        \hat f_{3s}  & = \frac{3}{2} \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        \hat f_{4sc} & = \frac{15}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        \hat f_{4s}  & = \frac{3}{2} \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \int_0^\pi \rmdx{\theta_2} \int_0^{2\pi} \rmdx{\phi} P(q^2, \theta_1, \theta_2, \phi) \hat{f}_i(\theta_1, \theta_2, \phi) = S_i(q^2)\,.
\end{equation}
